### 2022-07-16 preprint.tex
Deploying to gh-pages from @ graciellehigino/ms_range_interactions@79ba5448203e48bdc1cbd11a18cc1125b62a7850 🚀

Finding a species in a certain location is like finding an encrypted
message that traveled through time. It carries the species' evolutionary
history, migration patterns, as well as any direct and indirect effects
generated by other species (some of which we may not even know exist).
Ecologists have been trying to decode this message with progressively
more powerful tools, from their field notes to highly complex
computational algorithms. However, to succeed in this challenge it is
important to have the right clues in hand. There are many ways we can be
misled by data - or the lack of it: taxonomic errors (e.g., due to
updates in the taxonomy of a species), geographic inaccuracy (e.g.,
approximate coordinates or lack of documentation about their accuracy),
or sampling biases (e.g.~data clustered near roads or research centers)
(Ladle and Hortal 2013; Hortal et al. 2015; Poisot et al. 2021). One way
to identify - and potentially fix - these errors is to combine many
different pieces of information about the occurrence of a species, so
agreements and mismatches can emerge. Although previous studies have
combined different types of occurrence data to measure the accuracy of
datasets (Hurlbert and Jetz 2007; Hurlbert and White 2005;
\textbf{Ficetola2014EvaRoba?}), none have used different types of
information so far (i.e., ecological characteristics other than
geographical distribution). Here we suggest jointly analysing species
occurrence (range maps and point occurrences) and ecological
interactions to identify mismatches between datasets.

@@

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; \textbf{Ficetola2014EvaRoba?}). Adding ecological interaction data
to this comparison might help to elucidate where these (dis)agreements
are more likely to be true and which dataset better represent the actual
distribution of species. In this context, we elaborate a method that
allows us to refine distribution data (more precisely range maps) based
on interaction data, considering the basic assumption that predators can
only be present in regions where they are connected to at least one
herbivore - and thus indirectly connected to primary producers. We used
a Serengeti food web dataset (Baskerville et al. 2011) (which comprises
carnivores, herbivores, and plants from Tanzania) to demonstrate how a
mismatch between occurrence and interaction data can highlight
significant uncertainty areas in IUCN range maps. Finally, we add the
GBIF occurrence points for the Serengeti species to the investigation,
discuss the mechanisms that can lead to the lack of agreement between
data, and build from that a vision for the next steps, reinforcing the
importance of geographically explicit interaction data.

@@

\leavevmode\hypertarget{ref-Gotelli2010MacSig}{}%
---------. 2010. {``Macroecological Signals of Species Interactions in
the Danish Avifauna.''} \emph{Proceedings of the National Academy of
Sciences} 107 (11): 5030--35.
\url{https://doi.org/10.1073/pnas.0914089107}.

@@

\leavevmode\hypertarget{ref-Rondinini2006TraDif}{}%
---------. 2006. {``Tradeoffs of Different Types of Species Occurrence
Data for Use in Systematic Conservation Planning.''} \emph{Ecology
Letters} 9 (10): 1136--45.
\url{https://doi.org/10.1111/j.1461-0248.2006.00970.x}.

>>>>

Finding a species in a certain location is like finding an encrypted
message that traveled through time. It carries the species' evolutionary
history, migration patterns, as well as any direct and indirect effects
generated by other species (some of which we may not even know exist).
Ecologists have been trying to decode this message with progressively
more powerful tools, from their field notes to highly complex
computational algorithms. However, to succeed in this challenge it is
important to have the right clues in hand. There are many ways we can be
misled by data - or the lack of it: taxonomic errors (e.g., due to
updates in the taxonomy of a species), geographic inaccuracy (e.g.,
approximate coordinates or lack of documentation about their accuracy),
or sampling biases (e.g.~data clustered near roads or research centers)
(Ladle and Hortal 2013; Hortal et al. 2015; Poisot et al. 2021). One way
to identify - and potentially fix - these errors is to combine many
different pieces of information about the occurrence of a species, so
agreements and mismatches can emerge. Although previous studies have
combined different types of occurrence data to measure the accuracy of
datasets (Hurlbert and Jetz 2007; Hurlbert and White 2005; Ficetola et
al. 2014), none have used different types of information so far (i.e.,
ecological characteristics other than geographical distribution). Here
we suggest jointly analysing species occurrence (range maps and point
occurrences) and ecological interactions to identify mismatches between
datasets.

@@

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; Ficetola et al. 2014). Adding ecological interaction data to this
comparison might help to elucidate where these (dis)agreements are more
likely to be true and which dataset better represent the actual
distribution of species. In this context, we elaborate a method that
allows us to refine distribution data (more precisely range maps) based
on interaction data, considering the basic assumption that predators can
only be present in regions where they are connected to at least one
herbivore - and thus indirectly connected to primary producers. We used
a Serengeti food web dataset (Baskerville et al. 2011) (which comprises
carnivores, herbivores, and plants from Tanzania) to demonstrate how a
mismatch between occurrence and interaction data can highlight
significant uncertainty areas in IUCN range maps. Finally, we add the
GBIF occurrence points for the Serengeti species to the investigation,
discuss the mechanisms that can lead to the lack of agreement between
data, and build from that a vision for the next steps, reinforcing the
importance of geographically explicit interaction data.

@@

@@
